---
title: "Group 11 Final Report: Examining Risk Factors for Adverse COVID-19 Outcomes"
author: "Khadijah Abdallah, Leo Liu, Manju Ramakrishnan, Sarina Abrishamcar, Weiwei Wu"
date: "`r Sys.Date()`"
output: html_document
params:
  production: TRUE
---




```{r, set-chunk-opts, echo = FALSE, warning=FALSE}
library(knitr)
opts_chunk$set(
  echo = FALSE, warning = FALSE, message = FALSE
)
```

```{r, load-data}
here::i_am("reports/Group11_COVID_Final_Report.Rmd")
data <- readRDS(
  file = here::here("data/derived_data/data_clean.rds")
)
```

# Descriptive Analysis

## Tabular analysis

Characteristics for this dataset from Mexico for individuals with COVID-19 are displayed in the table below. 

Our study is examining the impact of diabetes on adverse COVID19 outcomes, including death and hospitalization. Table 1 presents results for individuals who were diagnosed with COVID-19 stratified by diabetes status. The results indicate that our study sample had slightly more males than females (53.4% vs. 46.6%, p=0.12). The overall average age was 44 years old. In terms of chronic diseases, about 18.6% of the study sample had obesity and 2.06% had chronic renal disease. Approximately 7% of the population were smokers. Individuals with diabetes were more likely to require hospitalization (56% vs. 23%, p <.001). They were also older (58 years vs. 41 years, p<.001), and more likely to have chronic diseases including obesity (28% vs. 17%, p<.001) and renal disease (7.3% vs. 1.1%, p<.001). Finally, they were more likely to be smokers (7.8% vs. 7.1%, p=.004) and to have died (32.3% vs. 10.1%, p<.001).   

```{r, table1}
table_one <- readRDS(
  file = here::here("output/table_1.rds")
)
table_one
```

## Data Distributions

Below we report on the data distributions for continuous and categorical variables. 

```{r, }

```

# Logistic Regression Analyses

## Model 1: Association Between Diabetes and Death From COVID-19

In this analysis, we assess the association between diabetes and death from COVID-19. We adjusted for sex, age, obesity status, presence of chronic renal disease, and smoking status. We conducted a logistic regression model using a generalized linear model with the binomial family and logit link. We identified a positive association between diabetes and death from COVID-19.The odds of death from COVID-19 among people with Diabetes is 1.85 (95% CI: 1.76, 1.95, P=<0.001) times the odds of death from COVID-19 among people without Diabetes.


```{r, fit-model1}
readRDS(
  file =here::here("output/Model1_table.RDS")
)

```

## Model 2: 

In this analysis, we examine the question of whether or not having diabetes increases one's risk for hospitalization. 

```{r, fit-model2}

```

