---
title: "Group 11 Final Report: Examining Risk Factors for Adverse COVID-19 Outcomes"
author: "Khadijah Abdallah, Leo Liu, Manju Ramakrishnan, Sarina Abrishamcar, Weiwei Wu"
date: "`r Sys.Date()`"
output: html_document
params:
  production: TRUE
---




```{r, set-chunk-opts, echo = FALSE, warning=FALSE}
library(knitr)
opts_chunk$set(
  echo = FALSE, warning = FALSE, message = FALSE
)
```

```{r, load-data}
here::i_am("reports/Group11_COVID_Final_Report.Rmd")
data <- readRDS(
  file = here::here("data/derived_data/data_clean.rds")
)
```

# Descriptive Analysis

## Tabular analysis

Characteristics for this dataset from Mexico for individuals with COVID-19 are displayed in the table below. 

Our study is examining the impact of diabetes on adverse COVID19 outcomes, including death and hospitalization. Table 1 presents results for individuals who were diagnosed with COVID-19 stratified by diabetes status. The results indicate that our study sample had slightly more males than females (53.4% vs. 46.6%, p=0.12). The overall average age was 44 years old. In terms of chronic diseases, about 18.6% of the study sample had obesity and 2.06% had chronic renal disease. Approximately 7% of the population were smokers. Individuals with diabetes were more likely to require hospitalization (56% vs. 23%, p <.001). They were also older (58 years vs. 41 years, p<.001), and more likely to have chronic diseases including obesity (28% vs. 17%, p<.001) and renal disease (7.3% vs. 1.1%, p<.001). Finally, they were more likely to be smokers (7.8% vs. 7.1%, p=.004) and to have died (32.3% vs. 10.1%, p<.001).   

```{r, table1}
table_one <- readRDS(
  file = here::here("output/table_1.rds")
)
table_one
```

## Data Distributions

Below we report on the data distributions for continuous and categorical variables. 

```{r, }

```

# Logistic Regression Analyses

## Model 1: Association Between Diabetes and Death From COVID-19

In this analysis, we assess the association between diabetes and death from COVID-19. We adjusted for sex, age, obesity status, presence of chronic renal disease, and smoking status. We conducted a logistic regression model using a generalized linear model with the binomial family and logit link. We identified a positive association between diabetes and death from COVID-19.The odds of death from COVID-19 among people with Diabetes is 1.85 (95% CI: 1.76, 1.95, P=<0.001) times the odds of death from COVID-19 among people without Diabetes.


```{r, fit-model1}
readRDS(
  file =here::here("output/Model1_table.RDS")
)

```

## Model 2: Association Between Diabetes and Hospitalizations due to Covid-19 

To further examine the association of having diabetes and the likelihood of hospitalizations among patients, we conducted a logistic regression analysis. The model included predictors such as diabetes, age, sex, obesity, tobacco use, and chronic renal  conditions. Our results revealed that individuals with diabetes had 2.10 times higher odds of being hospitalized compared to those without diabetes (95% CI: 2.00-2.19, p < 0.001). Age was also found to be a significant predictor, with each year increase in age associated with a 6% increase in the odds of hospitalization (OR: 1.06, 95% CI: 1.06-1.06, p < 0.001). Males exhibited 1.65 times higher odds of hospitalization compared to females (95% CI: 1.59-1.71, p < 0.001). Additionally, obesity was associated with a 1.36 times higher odds of hospitalization (95% CI: 1.31-1.42, p < 0.001). However, there was no significant association between tobacco use and hospitalization odds (OR: 1.01, 95% CI: 0.94-1.08, p = 0.8). Patients with renal chronic conditions had 3.41 times higher odds of hospitalization compared to those without such conditions (95% CI: 3.02-3.85, p < 0.001). These findings underscore the importance of managing chronic conditions like diabetes, as well as identifying the influence of factors like age, sex, obesity and renal health that increase the likelihood of hospitalization of COVID-19 patients. This would further assist in creating targeted management and intervention strategies to reduce complication and prevent hospitalizations.

```{r, fit-model2}

readRDS(
  file =here::here("output/Model2_table.RDS")
)

```

